---
title: "Health Research Assignment AP"
author: "Mohammad Raza Abidi (20323787) (AP) (COMP1013)"
date: "`r Sys.Date()`"
output:
pdf_document: default
html_document: default
---

------------------------------------------------------------------------

By including this statement, we the authors of this work, verify that:

• We hold a copy of this assignment that we can produce if the original is lost or damaged.

• We hereby certify that no part of this assignment/product has been copied from any other student’s work or from any other source except where due acknowledgement is made in the assignment.

• No part of this assignment/product has been written/produced for us by another person except where such collaboration has been authorised by the subject lecturer/tutor concerned.

• We are aware that this work may be reproduced and submitted to plagiarism detection software programs for the purpose of detecting possible plagiarism (which may retain a copy on its database for future plagiarism checking).

• We hereby certify that we have read and understand what the School of Computing, Engineering and Mathematics defines as minor and substantial breaches of misconduct as outlined in the learning guide for this unit. \*\*\*

```{r Question 1}
getwd()
setwd("/Users/mohammad/Downloads/")


#Q1. Write the code to analyse the distribution of COVID patients (confirmed or suspected) across counties. Write the code to investigate the distribution of the patients across age groups (e.g., 0-18, 19-35, 36-50, 51+). Visualise both the findings using the histogram. Explain your findings

# 1. Start By downloading and saving your CSV files on your operating system.

# 2. View your CSV files 

patients<-read.csv("patientsUG.csv")

encounters<-read.csv("encountersUG.csv")

conditions<-read.csv("conditionsUG.csv")

# 3. Use dplyr package to utilize the filter command for the data frames

library(dplyr)

# 4. View your data frame

View(conditions)

# 5.Filter data frame to identify how many suspected and confirmed Covid patients

SuspectedCOVID19Patients<-filter(conditions, CODE == 840544004)

ConfirmedCOVID19Patients<-filter(conditions, CODE == 840539006)

# 6. Merge both the data frames to easily identify the patients and their counties

PNC<- inner_join(patients, conditions, by = "X")


# 7. To make it easier to identify, write the codes of the patients

SuspectedNConfirmed<- PNC %>% filter(CODE == "840544004" | CODE == "840539006")


# 8. To figure out how many Covid 19 patients in each county, use the table command.

CountyNumber<- table(SuspectedNConfirmed$COUNTY)

print(CountyNumber)

# 9. Use ggplot2 for displaying the data as a histogram

library(ggplot2)

# The code written in Step Number 10 and 11 has been taken from Youtube Channel named The Data Bro as an inspiration and a reference for the assessment.
#Youtube Channel: https://www.youtube.com/watch?v=6e8d8CJG1a4&ab_channel=TheDataBro

# 10. Convert the table in to a data frame in order to plot the histogram

CountyNumberF <- as.data.frame(CountyNumber)

names(CountyNumberF) <- c("County", "Count")

# 11.  Create a histogram using the ggplot package

ggplot(CountyNumberF, aes(x = County, y = Count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Distribution of COVID Patients Across Counties",
       x = "County", y = "Number of COVID Patients") +
  theme(axis.text.x = element_text(angle = 80, vjust = 0.5, hjust=0.5))

#12. Therefore, we can see that amongst all the counties, the Middlesex County is the county with the most suspected and confirmed COVID patients, while all the others are just average.


#13. Install and load the lubridate package
library(lubridate)


#14. Convert BIRTHDATE to Date format

patients$BIRTHDATE <- as.Date(patients$BIRTHDATE, format = "%Y-%m-%d")

#15. Calculate age by finding interval between the BIRTHDATE and the current date.

patients$AGE <- as.integer(interval(patients$BIRTHDATE, Sys.Date()) / years(1))

#16. The age column should now be visible
head(patients)

#17. Make ages into Groups
patients$AGE <- cut(patients$AGE, 
                                     breaks = c(0, 18, 35, 50, Inf), 
                                    labels = c("0-18", "19-35", "36-50", "51+"))

#18. Count the patients in each age group
AgeCountInGroups<- table(patients$AGE)
AgeGroupCountF <- as.data.frame(AgeCountInGroups)
names(AgeGroupCountF) <- c("AgeGroup", "Patients")

#19. Use the package ggplot2 again in order to display the histogram

library(ggplot2)
ggplot(AgeGroupCountF, aes(x = AgeGroup, y = Patients)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Distribution of Patients Across Age Groups",
       x = "Age Group", y = "Number of Patients")

#20. By analyzing the results, I have come to the conclusion that out of all the age groups,the 51+ age group has the most patients with a number of 4677, while the other age groups have only a 200-300 gap. 

```

```{r Question 2}
#Q2. Filter those patients in the data set that have contracted COVID-19 or Suspected COVID-19; ; what are the top 10 most common conditions (symptoms) related to the patients? Do the conditions differ between genders? Provide a table to rank the top 10 conditions for male and female patients separately. Elaborate on the findings.


#1. Gather the 10 most common conditions for all patients

#2. Utilizing the PNC Data Frame, filter the Gender and Description column
invisible({
  PNC %>% 
    filter(!is.na(GENDER), !is.na(DESCRIPTION))
})

#3. Use count function to count the number of each symptoms
EveryCondition <- PNC %>% count(DESCRIPTION)

#4. Use the top_n to view the top rows and the 10 most common conditions 
TenMostCommonConditions<- EveryCondition %>% top_n(10, n)

#5. By looking at the table and analyzing, we can clearly see which top 10 conditions are the most common and relatable to the patients

#6. Now the data needs to be distributed by Gender

GenderNDescription <- inner_join(PNC, TenMostCommonConditions, by = "DESCRIPTION")

Top10MC <- GenderNDescription %>% filter(GENDER == "M")



```
